Presentation Title

### 2017-08-13 thesis/chapters/ch1-intro.tex
Give paper a structure
Style improvements in thesis

Sadly, as traditional blockchain systems began to gain popularity,
their limitations also became apparent.
Bitcoin has the infamous 7 transactions per second upper bound~\cite{vukolic2015quest}.
This is due to the fact that blocks are fixed to 1 MB,
and are only generated on average every 10 minutes.
Since every Bitcoin transaction is at least 250 bytes, it computes to about 7 transactions per second.
Due to this limitation, it is not uncommon to see a long backlog of about 20,000 transactions\footnote{\url{https://blockchain.info/unconfirmed-transactions/}}.
A few months ago the backlog even reached 100,000, which meant new transactions would take at least 11 hours to be written to the blockchain~\cite{bitcoinbacklog}.
This issue has plagued the Bitcoin community for some time and is the root cause for the block size debate, which some calls it a civil war~\cite{bitcoincivilwar}.
Parties that are for the increase in block size argue that a larger block would improve the transaction rate.
Parties against it argue that it would make mining more centralised because blocks take longer to propagate through the network.
It also requires a hard fork (not backwards compatible), which risks consensus failure and devaluation.
A recent empirical study by Croman et el.~\cite{croman2016scaling} has shown that increasing the block size may help.
But given the bandwidth and latency constraints,
it is not possible to have more than 758 transactions per second.
Even worse, it may give some miner an unfair lead over others.
Thus fundamental changes are necessary to run at the scale of centralised payment processors such as Visa,
which is in the order of tens of thousands transactions per second~\cite{visa}.

@@

Using the aforementioned insight,
we explore an alternative consensus model for blockchain systems where transactions themselves do not reach consensus,
but nevertheless verifiable at a later stage by any node in the network.
We call our protocol Checo, derived from \emph{che}ckpoint \emph{co}nsensus.
Informally, our system works as follows.
Every node stores its own blockchain and every block is one transaction.
We randomly select nodes in every round, the selected ones are called facilitators.
Facilitators reach consensus not on the individual transactions,
but on the state of every chain represented by a single digest, called a checkpoint block.
If a checkpoint block of some node is in consensus, 
then that node can proof to any other node that it holds a set of transactions that compute (form a chain) to the checkpoint block.
This immediately shows that those transactions are tamper-proof.
The key difference from related work is the independence of the transaction validation phase.
Prior systems such a Bitcoin and Ethereum entangled the consensus algorithm with the transaction validation, restricting performance.
By splitting this functionality we achieved horizontal scalability in the general case.

Our main contributions are the following.
\begin{itemize}
    \item We formally introduce a blockchain system---Checo.
        It uses individual hash chains and checkpoints on every node to achieve
        horizontally scalable for the first time.
    \item We formally analyse Checo to ensure correctness as defined in our architecture.
    \item We provide an implementation and then experiment with up to 1200 nodes,
        our results show strong evidence of horizontal scalability.
\end{itemize}

>>>>

Sadly, as traditional blockchain systems began to gain popularity,
their limitations also became apparent.
Bitcoin has the infamous 7 transactions per second upper bound~\cite{vukolic2015quest}.
This is due to the fact that blocks are fixed to 1 MB,
and are only generated on average every 10 minutes.
Since every Bitcoin transaction is at least 250 bytes, it computes to about 7 transactions per second.
Due to this limitation, it is not uncommon to see a long backlog of about 20,000 transactions\footnote{\url{https://blockchain.info/unconfirmed-transactions/}}.
A few months ago the backlog even reached 100,000, which meant new transactions would take at least 11 hours to be written to the blockchain~\cite{bitcoinbacklog}.
This issue has plagued the Bitcoin community for some time and is the root cause for the block size debate, which some calls it a civil war~\cite{bitcoincivilwar}.
Parties that are for the increase in block size argue that a larger block would improve the transaction rate.
Parties against it argue that it would make mining more centralised because blocks take longer to propagate through the network.
It also requires a hard (non-backwards compatible) fork,
which risks consensus failure and devaluation.
A recent empirical study by Croman et el.~\cite{croman2016scaling} has shown that increasing the block size or the block interval may help.
But given the bandwidth and latency constraints of today's network,
it is not possible to have more than 758 transactions per second if messages need to be propagated timely to 90\% of the nodes.
To achieve transaction rates higher than that, it will lead to unfairness for nodes that do not have a good network connection.
Thus fundamental changes are necessary to run at the scale of centralised payment processors such as Visa,
which is in the order of tens of thousands transactions per second~\cite{visa}.

@@

Using the aforementioned insight,
we explore an alternative consensus model for blockchain systems where transactions themselves do not reach consensus,
but nevertheless verifiable at a later stage by any node in the network.
We call our protocol \textsf{Checo}, derived from \emph{che}ckpoint \emph{co}nsensus.
Informally, our system works as follows.
Every node stores its own blockchain and every block is one transaction.
We randomly select nodes in every round, the selected ones are called facilitators.
Facilitators reach consensus not on the individual transactions,
but on the state of every chain represented by a single digest, called a checkpoint block.
If a checkpoint block of some node is in consensus, 
then that node can proof to any other node that it holds a set of transactions that compute (form a chain) to the checkpoint block.
This immediately shows that those transactions are tamper-proof.
The key difference from related work is the independence of the transaction validation phase.
Prior systems such a Bitcoin and Ethereum entangled the consensus algorithm with the transaction validation, restricting performance.
By splitting this functionality we achieved horizontal scalability in the general case.
That is, by adding new nodes into the network, the global transaction rate should increase proportionally.

Our main contributions are the following.
\begin{itemize}
    \item We formally introduce a blockchain system---\textsc{Checo}.
        It uses individual hash chains and checkpoints on every node to achieve
        horizontally scalable for the first time.
    \item We formally analyse \textsc{Checo} to ensure correctness as defined in our architecture.
    \item We provide an implementation and then experiment with up to 1200 nodes,
        our results show strong evidence of horizontal scalability.
\end{itemize}

### 2017-08-13 thesis/chapters/ch3-model.tex
Style improvements in thesis

Our consensus protocol runs in rounds as discussed in \Cref{sec:system-overview}.
Every round is identified by a round number $r$, which is incremented on every new round.
The consensus result is a tuple, i.e. $\C_r = \langle r, C \rangle$,
where $C$ is a set of CP blocks agreed by the facilitators of round $r$.
\end{definition}

@@

$\forall r \in \mathbb{N}$, the following properties must hold.
\begin{enumerate}
    \item \emph{Agreement}:
        If one correct node outputs a set of facilitators $\F_r$,
        then every node outputs $\F_r$
    \item \emph{Validity}:
        If any correct node outputs $\F_r$, then 
            \begin{enumerate}
                \item $|\C_r| \ge N - t$ and each CP in $\C_r$ belong to a different node\footnote{
                $\C_r$ is a tuple but we abuse the notation here by writing $|\C_r|$ to mean the number of CP blocks in the second element of $\C_r$.}.
                \item $\F_r$ must contain at least $n - t$ honest nodes and
                \item $|\F_r| = n$.
            \end{enumerate}
    \item \emph{Fairness}:
        Every node with a CP block in $\C_r$ should have an equal probability of becoming a member of $\F_r$.
    \item \emph{Termination}:
        Every correct node eventually outputs some $\F_r$.
\end{enumerate}
\end{definition}
These properties are similar to Byzantine consensus properties (which we describe next in \Cref{sec:acs-background}), but there are subtle differences.
Firstly, they are properties for every node in the network and not just the facilitators.
Secondly, they must be satisfied for all rounds because the whole system falls apart if one of the property cannot be satisfied in one of the rounds.

@@

Upon receiving $\C_{r + 1}$ and at least $n - f$ valid signatures by some node $u$, $u$ performs two tasks.
First, it creates a new CP block using $\textsf{new\_cp}(\C_{r + 1}, r + 1)$ using \Cref{alg:new-cp}.
Second, it computes the new facilitators using $\textsf{get\_facilitator}(\C_{r + 1}, n)$ using \Cref{alg:facilitator},
and updates its facilitator set to the result.
This concludes the consensus phase and brings us back to the beginning of the consensus phase.

@@

The TX protocol, shown in \Cref{alg:tx-proto}, is run by all nodes.
It is also known as True Halves, first described by Veldhuisen~\cite[Chapter~3.2]{truehalves}.
Nodes that wish to initiate a transaction calls $\textsf{new\_tx}(pk_v, m, txid)$ (\Cref{alg:new-tx}) with the intended counterparty $v$ identified by $pk_v$ and message $m$.
$txid$ should be a uniformly distributed random value, i.e. $txid \in_R \{0, 1\}^{256}$.
Then the initiator sends $\langle \texttt{tx\_req}, t_{u, h}\rangle$ to $v$.

@@

\subsection{Validity definition}
A transaction can be in one of three states in terms of validity---\emph{valid}, \emph{invalid} and \emph{unknown}.
Given a fragment $F_{v, j}$, the validity of the TX block $t_{u, i}$ and the agreed fragment of it is captured by the function $\textsf{get\_validity}(t_{u, i}, F_{u, i}, F_{v, j})$ (\Cref{alg:get-validity}).
Note that $t_{u, i}$ and $F_{u, i}$ are assumed to be valid,
otherwise the node calling the function would have no point of reference.
This is not difficult to achieve.
Typically the caller is $u$, so it knows its own TX block and the corresponding agreed fragment.
If the caller is not $u$, it can always query for the agreed fragment that contains the transaction of interest from $u$.

@@

Further, the caller of $\textsf{get\_validity}(t_{u, i}, F_{u, i}, F_{v, j})$ is not necessarily 
$u$\footnote{In practice, it often is because after completing the TX protocol the parties are incentivised to check that the counterparty ``did the right thing''.}
Any node $w$ may call $\textsf{get\_validity}(t_u{u, i}, F_{u, i}, F_{v, j})$ as long as $w$ has an the necessary input parameters.
$F_{u, i}$  and $t_{u, i}$ may be readily available if $w = u$.
Otherwise, $w$ could query $u$ with the \texttt{vd\_req} message and then obtain the result from \texttt{vd\_resp}.
This is the validation protocol which we describe next.

@@

\subsection{Optimising validation protocol using cached agreed fragments}
\label{sec:caching}
One more way to improve the efficiency of the validation protocol is to use a single agreed fragment to validate multiple transactions.
Concretely, upon receiving an agreed fragment from node $A$,
rather than validating a single transaction,
we attempt to validate all transactions with $A$ in the unknown state in that fragment.

For a node, the benefit of this technique is maximised when it only transacts with one other node.
In this case, the communication of one fragment is sufficient to validate all transactions in that fragment.
In the opposite extreme, if every transaction that the node makes is with another unique node,
then the caching mechanism would have no effect.

\subsection{Global fork detection}
The validation algorithm guarantees that there are no forks within a single agreed fragment.
This is sufficient for some applications such as proving the existence of some information.
However, for applications such as digital cash where every block depends on one or more previous blocks,
our scheme is not suitable.
For such applications, we need to guarantee that there are no forks from the genesis block leading up to the TX block of interest.

There are a variety of approaches to do global fork detection.
First and the easiest solution is to simply ask for the complete chain of the counterparty.
The verifier can be sure that there are no forks if the following holds.
\begin{enumerate}
    \item The hash pointers are correct.
    \item All the CP blocks are in consensus.
    \item The TX of interest is in the chian.
\end{enumerate}
We use this approach in our prior work on Implicit Consensus~\cite{implicitconsensus}.
It may sound inefficient at first, but nodes employ caching to minimise communication costs.

>>>>

Our consensus protocol runs in rounds as discussed in \Cref{sec:system-overview}.
Every round is identified by a round number $r$, which is incremented on every new round.
The consensus result is a tuple, i.e. 
$$\C_r = \langle r, C \rangle,$$
where $C$ is a set of CP blocks agreed by the facilitators of round $r$.
\end{definition}

@@

$\forall r \in \mathbb{N}$, the following properties must hold.
\begin{enumerate}
    \item \emph{Agreement}:
        If one correct node outputs a set of facilitators $\F_r$,
        then every node outputs $\F_r$
    \item \emph{Validity}:
        If any correct node outputs $\F_r$, then 
            \begin{enumerate}
                \item $|\C_r| \ge N - t$ and each CP in $\C_r$ belong to a different node\footnote{
                $\C_r$ is a tuple but we abuse the notation here by writing $|\C_r|$ to mean the number of CP blocks in the second element of $\C_r$.}.
                \item $\F_r$ must contain at least $n - t$ honest nodes and
                \item $|\F_r| = n$.
            \end{enumerate}
    \item \emph{Fairness}:
        Every node with a CP block in $\C_r$ should have an equal probability of becoming a member of $\F_r$.
    \item \emph{Termination}:
        Every correct node eventually outputs some $\F_r$.
\end{enumerate}
\end{definition}
These properties are similar to Byzantine consensus properties (which we describe next in \Cref{sec:acs-background}), but there are subtle differences.
Firstly, they are properties for every node in the network and not just the facilitators.
Secondly, they must be satisfied for all rounds,
because a failure in one round will lead to failures in sebsequent rounds.

@@

Upon receiving $\C_{r + 1}$ and at least $n - f$ valid signatures by some node $u$, $u$ performs two tasks.
First, it creates a new CP block using $\textsf{new\_cp}(\C_{r + 1}, r + 1)$ (\Cref{alg:new-cp}).
Second, it computes the new facilitators using $\textsf{get\_facilitator}(\C_{r + 1}, n)$ (\Cref{alg:facilitator}),
and updates its facilitator set to the result.
This concludes the consensus phase and brings us back to the state at the beginning of the consensus phase,
so a new round can be started.

@@

The TX protocol, shown in \Cref{alg:tx-proto}, is run by all nodes.
It is also known as True Halves, first described informally by Veldhuisen~\cite[Chapter~3.2]{truehalves}.
Nodes that wish to initiate a transaction calls $\textsf{new\_tx}(pk_v, m, txid)$ (\Cref{alg:new-tx}) with the intended counterparty $v$ identified by $pk_v$ and message $m$.
$txid$ should be a uniformly distributed random value, i.e. $txid \in_R \{0, 1\}^{256}$.
Then the initiator sends $\langle \texttt{tx\_req}, t_{u, h}\rangle$ to $v$.

@@

\subsection{Validity definition}
A transaction can be in one of three states in terms of validity---\emph{valid}, \emph{invalid} and \emph{unknown}.
Given a fragment $F_{v, j}$, the validity of the TX block $t_{u, i}$ with its corresponding fragment $F_{u, i}$ is captured by the function $\textsf{get\_validity}(t_{u, i}, F_{u, i}, F_{v, j})$ (\Cref{alg:get-validity}).
Note that $t_{u, i}$ and $F_{u, i}$ are assumed to be valid,
otherwise the node calling the function would have no point of reference.
This is not difficult to achieve,
because typically the caller is $u$, so it knows its own TX block and the corresponding agreed fragment.
If the caller is not $u$, it can always query for the agreed fragment that contains the transaction of interest from $u$.

@@

% Further, the caller of $\textsf{get\_validity}(t_{u, i}, F_{u, i}, F_{v, j})$ is not necessarily 
% $u$\footnote{In practice, it often is because after completing the TX protocol the parties are incentivised to check that the counterparty ``did the right thing''.}
% Any node $w$ may call $\textsf{get\_validity}(t_u{u, i}, F_{u, i}, F_{v, j})$ as long as $w$ has an the necessary input parameters.
% $F_{u, i}$  and $t_{u, i}$ may be readily available if $w = u$.
% Otherwise, $w$ could query $u$ with the \texttt{vd\_req} message and then obtain the result from \texttt{vd\_resp}.
% This is the validation protocol which we describe next.

@@

\subsection{Optimising validation protocol using cached agreed fragments}
\label{sec:caching}
One more way to improve the efficiency of the validation protocol is to use a single agreed fragment to validate multiple transactions.
Concretely, for node $A$, upon receiving an agreed fragment from node $B$,
rather than validating a single transaction,
$A$ attempts to validate all transactions performed with $B$ in the unknown state, in that fragment.

The benefit of this technique is maximised when a node only transacts with one other node.
In this case, the communication of one fragment is sufficient to validate all transactions in that fragment.
In the opposite extreme, if every transaction that the node makes is with another unique node,
then the caching mechanism would have no effect.

\subsection{Total fork detection}
The validation algorithm guarantees that there are no forks within a single agreed fragment.
This is sufficient for some applications such as proving the existence of some information.
However, for applications such as digital cash where every block depends on one or more previous blocks,
our scheme is not suitable.
For such applications, we need to guarantee that there are no forks from the genesis block leading up to the TX block of interest.

We offer two approaches to do total fork detection.
First and the easiest solution is to simply ask for the complete chain of the counterparty.
The verifier can be sure that there are no forks if the following holds.
\begin{enumerate}
    \item The hash pointers are correct.
    \item All the CP blocks are in consensus.
    \item The TX of interest is in the chian.
\end{enumerate}
We use this approach in our prior work on Implicit Consensus~\cite{implicitconsensus}.
Nodes employ caching to minimise communication costs, we call this effect spontaneous sharding.
